---
title: "CFB Stats Generic Scraper"
---

```{r}
#| label: setup
#| message: false
library(rvest)
library(dplyr)
library(janitor)
```

## Generic Scraper Function

Pass any URL path from [cfbstats.com/2025/national/index.html](https://cfbstats.com/2025/national/index.html)
along with a year. The path is the portion of the URL **after** the year segment.

For example, for this URL:

```
https://cfbstats.com/2025/leader/national/team/offense/split01/category09/sort01.html
```

The path argument would be:

```
/leader/national/team/offense/split01/category09/sort01.html
```

```{r}
#| label: scraper-function

#' Scrape any stat leaderboard from cfbstats.com
#'
#' @param path Character. The URL path after the year, starting with "/".
#'   Example: "/leader/national/team/offense/split01/category09/sort01.html"
#' @param year Integer. Season year (e.g., 2025). Available years: 2016–2025.
#' @return A data frame of the stats table with a `year` column appended.
scrape_cfbstats <- function(path, year) {
  # Strip leading slash if present so sprintf doesn't double up
  path <- sub("^/", "", path)

  url <- sprintf("https://cfbstats.com/%d/%s", year, path)

  page <- tryCatch(
    read_html(url),
    error = function(e) stop("Failed to fetch: ", url, "\n", e$message)
  )

  tbl_node <- html_element(page, "table.leaders")

  if (is.na(tbl_node)) {
    stop("No table with class 'leaders' found at: ", url)
  }

  tbl_node |>
    html_table(header = TRUE) |>
    clean_names() |>            # snake_case column names via janitor
    mutate(year = year, .before = 1)
}
```

## Example Usage

### Single page + year

```{r}
#| label: example-single

scoring_offense <- scrape_cfbstats(
  path = "/leader/national/team/offense/split01/category09/sort01.html",
  year = 2025
)

head(scoring_offense)
```

### Loop over multiple years

```{r}
#| label: example-multi-year

rushing_defense_multi <- lapply(2022:2025, function(yr) {
  Sys.sleep(0.5)  # be polite to the server
  scrape_cfbstats(
    path = "/leader/national/team/defense/split01/category01/sort01.html",
    year = yr
  )
}) |>
  bind_rows()

glimpse(rushing_defense_multi)
```

### Player stats example

```{r}
#| label: example-player

passing_leaders <- scrape_cfbstats(
  path = "/leader/national/player/split01/category02/sort01.html",
  year = 2025
)

head(passing_leaders)
```

### Try another

I tried to use copilot to complete but it didn't get the url right, but I fixed it manually.

```{r}
total_offense <- scrape_cfbstats(
  path = "leader/national/team/offense/split01/category10/sort01.html",
  year = 2025
)

# show the result
head(total_offense)
```

